\documentclass{oblivoir}

\usepackage{amsthm}
\usepackage{thmtools}

\usepackage{url}

%\usepackage{ansform}
\usepackage{ikps}

\usepackage{graphicx}
%https://www.overleaf.com/learn/latex/Inserting_Images

\usepackage{listings}
\usepackage{xcolor}
\declaretheoremstyle[% spaceabove=6pt,spacebelow=6pt, headfont=\color{MainColorOne}\sffamily\bfseries, notefont=\mdseries, notebraces={[}{]}, bodyfont=\normalfont,
headpunct={},
postheadspace=1em,
%qed=▣,
]{maintheorem}

\declaretheorem[%
name=정의,
style=maintheorem,
numberwithin=section, shaded={%bgcolor=MainColorThree!20,
margin=.5em}]{dfn}
% \begin{dfn}[]
% \end{dfn}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
%https://www.overleaf.com/learn/latex/Theorems_and_proofs

\definecolor{mGreen}{rgb}{0,0.6,0}
\definecolor{mGray}{rgb}{0.5,0.5,0.5}
\definecolor{mPurple}{rgb}{0.58,0,0.82}
\definecolor{backgroundColour}{rgb}{0.95,0.95,0.92}
%https://tex.stackexchange.com/questions/348651/c-code-to-add-in-the-document
\lstdefinestyle{CStyle}{
    backgroundcolor=\color{backgroundColour},   
    commentstyle=\color{mGreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{mGray},
    stringstyle=\color{mPurple},
    basicstyle=\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2,
    language=C
}

\title{Introduction to Quick Sort}
\author{EUnS}


%https://tex.stackexchange.com/questions/246/when-should-i-use-input-vs-include
\begin{document}
    

\maketitle


\tableofcontents


\include{./section/section1}

\include{./section/section2}

\include{./section/section3}

\include{./section/section4}

\include{./section/section5}

\include{./section/section6}

\include{./section/section7}


\listoffigures

\listoftables

\newpage
% https://www.overleaf.com/learn/latex/Sections_and_chapters
\addcontentsline{toc}{section}{참고 자료}
\section*{참고 자료}
\url{https://en.wikipedia.org/wiki/Quicksort} \\
\url{https://cs.stackexchange.com/questions/11458/ quicksort-partitioning-hoare-vs-lomuto} \\
\url{https://www.acmicpc.net/blog/view/58} \\
\url{https://www.youtube.com/watch?v=hq4dpyuX4Uw&list=PL52K_8WQO5oUuH06MLOrah4h05TZ4n38l&index=11} \\
\url{https://en.wikipedia.org/wiki/Recursion_(computer_science)} \\
% recursion
\url{https://algs4.tistory.com/45} \\
% h vs l testing
\url{https://arxiv.org/pdf/1905.00118.pdf} \\
\url{http://rohitja.in/lomuto-hoare-partitioning.html} \\


\url{https://www.cs.princeton.edu/~rs/talks/QuicksortIsOptimal.pdf}
\url{https://en.wikipedia.org/wiki/Dutch_national_flag_problem}\\
% 3 way partition
\url{https://en.wikipedia.org/wiki/Introsort}\\
%introsort
\url{https://www.geeksforgeeks.org/internal-details-of-stdsort-in-c/}
%std::sort simple code
\url{https://stackoverflow.com/questions/44441876/quick-sort-using-stack-in-c}
%quicksort with stack 

\url{http://fpl.cs.depaul.edu/jriely/ds1/extras/lectures/23Quicksort.pdf}

\begin{thebibliography}{}
    \bibitem{reference1}
    Thomas H. Cormen, Charles E. Leiserson, Ronald L. Rivest, and Clifford Stein. Introduction to Algorithms, Second Edition. MIT Press and McGraw-Hill, 2001. ISBN 0-262-03293-7.
    
    \bibitem{reference2}
    Patterson, David A./ Hennessy, John L. Computer Organization and Design, Fourth Edition: The Hardware/Software Interface (The Morgan Kaufmann Series in Computer Architecture and Design). 2014. ISBN-13 9788994961897, ISBN-10 8994961895

    \bibitem{reference3}
    Sedgewick, R. (1978). "Implementing Quicksort programs". Comm. ACM. 21 (10): 847–857. doi:10.1145/359619.359631.

    \bibitem{reference4}
    Robert Sedgewick, Kevin Wayne Algorithms, 4th Edition, 2001 ISBN-13: 978-0321573513

\end{thebibliography}
\end{document}

% testing
% std::sort