
% https://www.overleaf.com/learn/latex/Sections_and_chapters
\addcontentsline{toc}{section}{참고 자료}
\section*{참고 자료}
\url{https://en.wikipedia.org/wiki/Quicksort} \\
\url{https://cs.stackexchange.com/questions/11458/ quicksort-partitioning-hoare-vs-lomuto} \\
\url{https://www.acmicpc.net/blog/view/58} \\
\url{https://www.youtube.com/watch?v=hq4dpyuX4Uw&list=PL52K_8WQO5oUuH06MLOrah4h05TZ4n38l&index=11} \\
\url{https://en.wikipedia.org/wiki/Recursion_(computer_science)} \\
% recursion
\url{https://algs4.tistory.com/45} \\
% h vs l testing
\url{https://arxiv.org/pdf/1905.00118.pdf} \\
\url{http://rohitja.in/lomuto-hoare-partitioning.html} \\


\url{https://www.cs.princeton.edu/~rs/talks/QuicksortIsOptimal.pdf}
\url{https://en.wikipedia.org/wiki/Dutch_national_flag_problem}\\
% 3 way partition
\url{https://en.wikipedia.org/wiki/Introsort}\\
%introsort
\url{https://www.geeksforgeeks.org/internal-details-of-stdsort-in-c/}
%std::sort simple code
\url{https://stackoverflow.com/questions/44441876/quick-sort-using-stack-in-c}
%quicksort with stack 

\url{http://fpl.cs.depaul.edu/jriely/ds1/extras/lectures/23Quicksort.pdf}

\begin{thebibliography}{}
    \bibitem{reference1}
    Thomas H. Cormen, Charles E. Leiserson, Ronald L. Rivest, and Clifford Stein. Introduction to Algorithms, Second Edition. MIT Press and McGraw-Hill, 2001. ISBN 0-262-03293-7.
    
    \bibitem{reference2}
    Patterson, David A./ Hennessy, John L. Computer Organization and Design, Fourth Edition: The Hardware/Software Interface (The Morgan Kaufmann Series in Computer Architecture and Design). 2014. ISBN-13 9788994961897, ISBN-10 8994961895

    \bibitem{reference3}
    Sedgewick, R. (1978). "Implementing Quicksort programs". Comm. ACM. 21 (10): 847–857. doi:10.1145/359619.359631.

    \bibitem{reference4}
    Robert Sedgewick, Kevin Wayne Algorithms, 4th Edition, 2001 ISBN-13: 978-0321573513

\end{thebibliography}